---
title: "The Best and Worst Places to Remote Work"
author: "scurry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(tigris_use_cache = TRUE)
setwd("C:/LocalDocuments/MUSA6310c/DINK_Math")

library(dplyr)
library(ggplot2)
library(knitr)
library(scales)
library(sf)
library(tidycensus)
library(tigris)

us <- states() %>% filter(as.numeric(REGION) <= 3)
water <- st_as_sfc(st_bbox(us))
pacounty <- counties(state = 42) %>% select(NAME)

acs_variable_list.2022 <- load_variables(2022, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

#pawage <- read.csv("LivingWage_PA2024.csv") %>% select(COUNTY, X2A2W0C) %>% rename("NAME" = "COUNTY")
#pawage <- merge(pacounty, pawage, by = "NAME", all.x = TRUE)

#pals_24 <- read.csv("LivingSalary_PA2024.csv")

#pals_2a1w0c_24 <- pals_24 %>% select(COUNTY, X2A1W0C) %>% rename("NAME" = "COUNTY") #individual income
#pals_2a2w0c_24 <- pals_24 %>% select(COUNTY, X2A2W0C) %>% rename("NAME" = "COUNTY") #household income
#pals_2a2w1c_24 <- pals_24 %>% select(COUNTY, X2A2W1C) %>% rename("NAME" = "COUNTY") #household income

#pals_2a1w0c_24 <- merge(pacounty, pasal, by = "NAME", all.x = TRUE)

msa_mhi_22 <- get_acs(geography = "metropolitan statistical area/micropolitan statistical area",
                     year = 2022, 
                     variables = "B19013_001",#median household income
                     geometry = TRUE
) 

msa_mhi_22 <- msa_mhi_22 %>% 
  arrange(estimate)
```

If you work remotely out of Philadelphia, where in the Northeastern region can you
live comfortably? 

In 2022, the Median Household Income in the Metropolitan Statistical Area of 
Philadelphia, PA was $85,555. Compared to the MHI of other MSAs in the United 
States,

```{r msa_mhi_data, message=FALSE, warning=FALSE, include=FALSE}

phl_mhi_22 <- msa_mhi_22 %>%
  filter(GEOID == 37980)

```


```{r msa_mhi_chart, echo=FALSE, message=FALSE, warning=FALSE}

ggplot() +
  geom_histogram(data = msa_mhi_22, aes(x = estimate), fill = "darkolivegreen4", color = "darkolivegreen") +
  scale_x_continuous(labels = scales::label_dollar()) +
  geom_vline(data = phl_mhi_22, aes(xintercept = estimate), color = "goldenrod", linetype = "dashed",
             size = 1) +
  geom_text(data = phl_mhi_22, aes(x = estimate + 3000, y = 100, label = "Philadlephia MSA"),
            color = "goldenrod", angle = 90) +
  labs(title = "2022 Median Household Income",
       subtitle = "US Metropolitan Statistical Areas",
       x = "Median Household Income",
       y = " ",
#       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  theme_minimal()

```

The Philadelphia MSA's MHI is in the 93rd percentile of all MSAs. Securing a
remote work job that was based in Philadelphia allows for a large amount of 
mobility. Generally, across the Northeast, a household earning that annual 
income could rent in most Counties without risk of being cost-burdened. 

A cost-burdened household pays more than 30% of its income on rent and utilities. 

The Cost of Living per County was calculated as the annual salary needed to spend 
30% on gross rent, meaning that the household would not be cost-burdened. 
In the maps below, Comfortability refers to the ratio of the Remote Work Base 
Location to the Cost of Living in the County.

```{r col_data, message=FALSE, warning=FALSE, include=FALSE}

ec_mgr_22 <- get_acs(geography = "county",
                     year = 2022, 
                     variables = "B25064_001",#GROSS RENT
                     geometry = TRUE,
                     state = c("FL", "GA", "SC", "NC", "TN",
                               "VA", "WV", "KY", "IN", "OH",
                               "MI", "MD", "DC", "PA", "NJ",
                               "NY", "CT", "RI", "MA", "NH",
                               "VT", "ME", "DE")
)

ec_msmoc_22 <- get_acs(geography = "county",
                     year = 2022, 
                     variables = "B25094_001",#SMOC
                     geometry = TRUE,
                     state = c("FL", "GA", "SC", "NC", "TN",
                               "VA", "WV", "KY", "IN", "OH",
                               "MI", "MD", "DC", "PA", "NJ",
                               "NY", "CT", "RI", "MA", "NH",
                               "VT", "ME", "DE")
)

ec_lrs_22 <- ec_mgr_22 %>%
  select(-moe) %>%
  mutate(livrsal = (estimate/.30)*12,
         pc_phl_mhi = (85555/livrsal)) %>%
  mutate(f_phl_mhi = case_when(pc_phl_mhi >= 2.0 ~ "Twice As Comfortable",
                               pc_phl_mhi < 2.0 & pc_phl_mhi >= 1.5 ~ "Nearly Twice as Comfortable",
                               pc_phl_mhi < 1.5 & pc_phl_mhi > 1.0 ~ "More than Comfortable",
                               pc_phl_mhi == 1.0 ~ "Equally As Comfortable",
                               pc_phl_mhi < 1.0 & pc_phl_mhi >= 0.8 ~ "Uncomfortable",
                               pc_phl_mhi < 0.8 & pc_phl_mhi >= 0.5 ~ "Moderately Uncomfortable",
                               pc_phl_mhi < 0.5 & pc_phl_mhi >= 0.3 ~ "Very Uncomfortable",
                               pc_phl_mhi < 0.3 ~ "Extremely Uncomfortable"
  ))

ec_los_22 <- ec_msmoc_22 %>%
  select(-moe) %>%
  mutate(livosal = (estimate/.30)*12,
         pc_phl_mhi = (85555/livosal)) %>%
  mutate(f_phl_mhi = case_when(pc_phl_mhi >= 2.0 ~ "Twice As Comfortable",
                               pc_phl_mhi < 2.0 & pc_phl_mhi >= 1.5 ~ "Nearly Twice as Comfortable",
                               pc_phl_mhi < 1.5 & pc_phl_mhi > 1.0 ~ "More than Comfortable",
                               pc_phl_mhi == 1.0 ~ "Equally As Comfortable",
                               pc_phl_mhi < 1.0 & pc_phl_mhi >= 0.8 ~ "Uncomfortable",
                               pc_phl_mhi < 0.8 & pc_phl_mhi >= 0.5 ~ "Moderately Uncomfortable",
                               pc_phl_mhi < 0.5 & pc_phl_mhi >= 0.3 ~ "Very Uncomfortable",
                               pc_phl_mhi < 0.3 ~ "Extremely Uncomfortable"
  ))

ec_lrs_22$f_phl_mhi <- factor(ec_lrs_22$f_phl_mhi, levels = c("Twice As Comfortable",
                                                              "Nearly Twice as Comfortable",
                                                              "More than Comfortable",
                                                              "Equally As Comfortable",
                                                              "Uncomfortable",
                                                              "Moderately Uncomfortable",
                                                              "Very Uncomfortable",
                                                              "Extremely Uncomfortable"))

ec_los_22$f_phl_mhi <- factor(ec_lrs_22$f_phl_mhi, levels = c("Twice As Comfortable",
                                                              "Nearly Twice as Comfortable",
                                                              "More than Comfortable",
                                                              "Equally As Comfortable",
                                                              "Uncomfortable",
                                                              "Moderately Uncomfortable",
                                                              "Very Uncomfortable",
                                                              "Extremely Uncomfortable"))

```

```{r col_map, echo=FALSE, message=FALSE, warning=FALSE}

ggplot() +
  geom_sf(data = water, fill = "steelblue") +
  geom_sf(data = ec_lrs_22, aes(fill = f_phl_mhi), color = "transparent") +
  scale_fill_brewer(palette = "RdYlGn", na.value = "gray", direction = -1) +
  xlim(-80, -65) +
  ylim(38, 45) +
  geom_sf(data = us, fill = "transparent", color = "steelblue", size = 8) +
  labs(title = "2022 Remote Work Bargains for Renters",
       subtitle = "Northeastern Region",
       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  theme_void() +
  theme(legend.position = "right")

ggplot() +
  geom_sf(data = water, fill = "steelblue") +
  geom_sf(data = ec_lrs_22, aes(fill = f_phl_mhi), color = "transparent") +
  scale_fill_brewer(palette = "RdYlGn", direction = -1) +
  xlim(-81, -74.5) +
  ylim(39.5, 42.5) +
  geom_sf(data = us, fill = "transparent", color = "steelblue", size = 8) +
  labs(title = "2022 Remote Work Bargains for Renters",
       subtitle = "Pennsylvania",
       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  theme_void() +
  theme(legend.position = "right")

```

Syracuse, NY 75th percentile
Brunswick, GA 50th percentile
Goldsboro, NC 25th percentile